Optimizing Abstract$^2$ Machines

### 2012-10-14 paper/paper.tex
Definition for compilation.
Code for generic fixpoint and widening.

\documentclass{llncs}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xltxtra}
\setmonofont[Scale=MatchLowercase]{DejaVu Sans Mono}
%
\input{preamble}
\begin{document}
\title{Optimizing Abstract$^2$ Machines}
%
\frontmatter          % for the preliminaries
%
\author{J.~Ian Johnson\inst{1} \and Matthew Might\inst{2} \and David Van Horn\inst{1}}
\institute{Northeastern University \and University of Utah}

@@

@@

@@

@@

>>>>

\documentclass{llncs}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{xltxtra}
\setmonofont[Scale=MatchLowercase]{DejaVu Sans Mono}
%
\input{preamble}
\begin{document}
\title{Optimizing Abstract$^2$ Machines}
%
\frontmatter          % for the preliminaries
%
\author{J.~Ian Johnson\inst{1} \and Matthew Might\inst{2} \and David Van Horn\inst{1}}
\institute{Northeastern University \and University of Utah}

@@

\subsection{Generic fixpoint calculator}

We compute the semantics of a program by iterating the state
transition relation until a fixed point in the reachable states is
reached.

\begin{verbatim}
;; appl : (∀ (X) ((X -> (Setof X)) -> ((Setof X) -> (Setof X))))
(define ((appl f) s)
  (for/fold ([i (set)])
    ([x (in-set s)])
    (set-union i (f x))))

;; Calculate fixpoint of (appl f).
;; fix : (∀ (X) ((X -> (Setof X)) (Setof X) -> (Setof X)))
(define (fix f s)
  (let loop ((accum (set)) (front s))
    (if (set-empty? front)
        accum
        (let ((new-front ((appl f) front)))
          (loop (set-union accum front)
                (set-subtract new-front accum))))))
\end{verbatim}

@@

\begin{verbatim}
;; State^ = (cons (Set Conf) Store)

;; (State -> Setof State) -> State^ -> { State^ }
(define ((wide-step step) state)
  (match state
    [(cons cs σ)
     (define ss ((appl step)
                 (for/set ([c cs]) (c->s c σ))))
     (set (cons (for/set ([s ss]) (s->c s))
                (join-stores ss)))]))
\end{verbatim}

@@

\begin{figure}
\begin{align*}
\compile{\svar\mvar} &= \lambda(\menv,\msto,\mkont) .\co{\mkont,\mval,\msto} \text{ where }\mval\in\msto(\menv(\mvar))
\\
\compile{\slit\mlit} &= \lambda(\menv,\msto,\mkont) .
\co{\mkont,\mlit,\msto}
\\
\compile{\slam\mvar\mexp} &= \lambda(\menv,\msto,\mkont) .
\co{\mkont,\clos{\mvar,\compile\mexp,\menv},\msto}
\\
\compile{\sapp[^\mlab]{\mexpi0}{\mexpi1}} &= \lambda^\mcntr(\menv,\msto,\mkont) .
\compile{\mexpi0}^\mcntr(\menv,\msto',\kar[_\mlab^\mcntr]{\compile{\mexpi1},\menv,\maddr})
\text{ where }\maddr,\msto' = \mathit{push}^\mcntr_\mlab(\msto,\mkont)
\\
\compile{\sif[^\mlab]{\mexpi0}{\mexpi1}{\mexpi2}} &= \lambda^\mcntr(\menv,\msto,\mkont) .
\compile{\mexpi0}^\delta(\menv,\msto',\kif[^\mcntr]{\compile{\mexpi1},\compile{\mexpi2},\menv,\maddr})
\text{ where }\maddr,\msto' = \mathit{push}_\mlab^\mcntr(\msto,\mkont)
\end{align*}
\caption{Compilation}
\end{figure}

\begin{figure}
\begin{align*}
\mathit{eval}(\mexp) &= \{ \mstate\ |\ \compile{\mexp}(\epsilon,\varnothing,\varnothing,\kmt) \multimachstep \mstate \} \text{ where }
\\[2mm]
%% CONTINUE
\co{\kmt,\mval,\msto} &\machstep
\ans{\msto,\mval}
\\
\co{\kar[^\mcntr_\mlab]{\mcomp,\menv,\maddr},\mval,\msto} & \machstep
\mcomp^\mcntr(\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr})
\\
\co{\kfn[^\mcntr_\mlab]{{\mvalx{u}},\maddr},\mval,\msto} & \machstep
\ap[^\mcntr_\mlab]{\mval,\mvalx{u},\mkont,\msto}
\text{ where }\mkont \in \msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\strue,\msto} & \machstep
\mcompi0^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\sfalse,\msto} & \machstep
\mcompi1^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\[2mm]
%% APPLY
\ap[^\mcntr_\mlab]{\clos{\mvar,\mcomp,\menv},\mval,\msto,\mkont} & \machstep
\mcomp^{\mcntr'}(\menv',\msto',\mkont)
\text{ where }\menv',\msto',\mcntr' = \mathit{bind\,}^{ \mcntr}_\mlab(\msto,\mvar,\mval)
\\
\ap{\mop,\mval,\msto,\mkont} & \machstep
\co{\mkont,\mval',\msto}
\text{ where }\mkont\in\msto(\maddr)
\text{ and } \mval'\in\interpdelta(\mop,\mval)
\end{align*}
\caption{Abstract$^2$ machine for compiled ISWIM}
\label{fig:caam}
\end{figure}

@@

Boucher and Feeley, CC'96 introduce the idea of abstract compilation
by closure generation.  (Feeley and Lapalme.)
